Assessing reliability of an online behavioural laterality battery: a pre-registered

### 2020-08-04 online_laterality_reliability.Rmd
comments in file

The basic demographics questionnaire asked participants to report age, gender, years in education, and whether they were bilingual. We also askedquestions about handedness and footedness. For handedness we asked: which is your dominant hand (i.e. which hand do you prefer to use for tasks such as writing, cutting, and catching a ball)? For footedness we asked: Pwhich foot do you normally use to step up on a ladder/step? For both questions, participants could answer "left", "right", or "no preference".

@@

@@

  LItable[i,2]<-length(temp)
  LItable[i,3]<-round(mean(temp),2)
  LItable[i,4]<-round(sd(temp),2)
  LItable[i,5]<-round(t.test(temp)$statistic,2)
  pstring <- round(wilcox.test(temp)$p.value,3) # use wilcoxon
  ifelse(pstring<.001,pstring<-'<0.001',pstring<-char(pstring))

@@

The dichotic listening and chimeric faces tasks had test-retest reliability above *r*= .75, and so we investigated whether the tasks could be made more efficient by reducing the number of trials while keeping test-retest reliability above our criterion level of *r*= .65. Hence, we recalculated LIs for the dichotic listening and chimeric faces tasks using 1 to n trials, where n is the total number of trials on each task. The absolute test-retest correlation coefficients are shown in Figure 5. A minimum of 60 trials was required to achieve a test-retest correlation of *r*= .65 using the dichotic listening task. For the chimeric faces task, a minimum of 25 trials is necessary to achieve a test-retest reliability of r= .65.

@@

@@

Figure S1 shows scatterplots indicating scores on left and right sides for each task, with handedness colour-coded. For each task, mean for each side is shown as vertical and horizontal dotted lines, and equal performance on the two sides corresponds to the sloping red dotted line: thus an impression of whether the task is lateralised can be obtained by considering whether points cluster around the sloping lines symmetrically.

@@

A preliminary impression of the data are shown in the Online Supplemental Materials.  For the rhyme decision task only correct responses were used for computing mean log RTs (shown transformed back to original ms scale in the plot). For dichotic listening and chimeric faces a correct response can be given from either side; it follows that there will be a negative correlation between the two sides reflecting the fact that the number of selections on the left will be inversely related to the number of correct responses on the right. For these tasks it is evident by inspection that most responses for dichotic listening cluster above the dotted line, indicating a right ear advantage, whereas most responses on chimeric faces cluster below the dotted line, indicating a left hemi-face advantage. Finally, we show the number of correct sequences with each hand on the finger tapping task. Here, a strong impact of handedness is evident on inspection, with left-handers generally more accurate with the left hand, and right-handers more accurate with the right hand.

>>>>

The basic demographics questionnaire asked participants to report age, gender, years in education, and whether they were bilingual. We also askedquestions about handedness and footedness. For handedness we asked: which is your dominant hand (i.e. which hand do you prefer to use for tasks such as writing, cutting, and catching a ball)? For footedness we asked: Pwhich foot do you normally use to step up on a ladder/step? For both questions, participants could answer "left", "right", or "no preference".<!---ZW: what options were they given? Left / right / ambidextrous? A reviewer asked me about this recently--->

@@

<!---ZW: This bit confused me.

I thought that the hypothesis was that extreme left handers were more likely to be atypical on any given task, but the test looks at consistency of laterality across pairs of task. Isn’t that testing a different hypothesis?

Both are interesting, but I think the first one needs to be tested too. E.g., are extreme left handers more likely than right handers to be atypically lateralized on Rhyme Decision / Chimeric Faces / Dichotic Listening?
--->

@@

  LItable[i,2]<-length(temp)
  LItable[i,3]<-round(mean(temp),2)
  LItable[i,4]<-round(sd(temp),2)
  LItable[i,5]<-round(wilcox.test(temp)$statistic,2)
  pstring <- round(wilcox.test(temp)$p.value,3) # use wilcoxon
  ifelse(pstring<.001,pstring<-'<0.001',pstring<-char(pstring))

@@

The dichotic listening and chimeric faces tasks had test-retest reliability above *r*= .75, and so we investigated whether the tasks could be made more efficient by reducing the number of trials while keeping test-retest reliability above our criterion level of *r*= .65. Hence, we recalculated LIs for the dichotic listening and chimeric faces tasks using 1 to n trials, where n is the total number of trials on each task. The absolute test-retest correlation coefficients are shown in Figure 5. A minimum of 60 trials was required to achieve a test-retest correlation of *r*= .65 using the dichotic listening task. For the chimeric faces task, a minimum of 25 trials is necessary to achieve a test-retest reliability of r= .65. 
<!--- I think this section is useful for us, but maybe not really for others. Is it just diluting the main messages of the paper? --->

@@

Woodhead Z. V. J., Bradshaw A. R., Wilson A. C., Thompson P. A. & Bishop D. V. M. (2019). Testing the unitary theory of language lateralization using functional transcranial Doppler sonography in adults. *Royal Society Open Science*, *6*, 181801.

@@

Figure S1 shows scatterplots indicating scores on left and right sides for each task, with self-reported handedness (left vs right) colour-coded. For each task, the mean for each side is shown as vertical and horizontal dotted lines, and equal performance on the two sides corresponds to the sloping red dotted line: thus an impression of whether the task is lateralised can be obtained by considering whether points cluster around the sloping lines symmetrically.

@@

For the rhyme decision task only correct responses were used for computing mean log RTs (shown transformed back to original ms scale in the plot). For dichotic listening and chimeric faces a correct response can be given from either side; it follows that there will be a negative correlation between the two sides reflecting the fact that the number of selections on the left will be inversely related to the number of correct responses on the right. For these tasks it is evident by inspection that most responses for dichotic listening cluster above the dotted line, indicating a right ear advantage, whereas most responses on chimeric faces cluster below the dotted line, indicating a left hemi-face advantage. Finally, we show the number of correct sequences with each hand on the finger tapping task. Here, a strong impact of handedness is evident on inspection, with left-handers generally more accurate with the left hand, and right-handers more accurate with the right hand.
